---
title: "Regression modeling with R"
subtitle: "Setup notebook"
author: "Karla Lindquist"
date: "October 7, 2021"
output: html_notebook
---

### Create an RStudio project 

I strongly recommend downloading the complete project from the [CLE page](http://tiny.ucsf.edu/dsiregr) or [GitHub](https://github.com/UCSF-DSI/RegR) so that you don't have to worry about working directories. This is the main benefit of using RStudio projects - wherever you save them, everything will be relative to the .Rproj file.  

If you want to creat an RStudio project from scratch, instructions are as follows:  
1. Go to the RStudio main menu, select __File -> New Project__.  
2. Select __New Directory__
3. Select __Empty Project__
4. Give the project an informative __Directory Name__ (this will be created for you). Use __Browse__ to choose the location for this directory.   
_Optional_: check the "Create git repository" box if you know how to use git for version control.
5. Click __Create Project__ button. This creates a new project directory, as well as a .Rproj file within it. 

Within the project folder that you just created, create the following directories: 

- Data (save hrs_analyze.csv file from the Datasets section of the CLE page here)
- Notebooks (if using R markdown)
- Scripts (if creating new project from scratch) 
- Results (optional)

</br>

### Start an R script


You can start a new R scipt using the RStudio drop-down menu that has a little page icon with a green **+** symbol (top left).

_Best practice tip_:  At the top of your script, write comments that indicate:

```{r}
## Purpose: save commands from Regression Modeling with R class
## Author: Karla Lindquist (karla.lindquist@ucsf.edu)
## Updated: October 7, 2021
```


You may also want to pre-define sections in your script. Using the `####` after a comment will create a navigatable menu in RStudio (see bottom of script window).  

```{r}
## Read data ####

## Liner regression ####

## Logistic regression ####

## Survival analysis ####

```

</br>

### Install and Load Required Packages  

Next, install and load the packages that we will be using in all sections.

```{r}
pkgs <- c("dplyr","ggplot2","jtools","lmtest","naniar","survival","gplots","survminer","ggstance","broom.mixed") 
## note: some may have issues with installing a few of these packages ... if you have trouble, please feel free to email karla.lindquist@ucsf.edu 

install.packages(pkgs, update = TRUE, ask = FALSE) 

for(package in pkgs) {
  do.call("library", list(package))
}
```

*Note: this will update packages if they already exist in your system. If asked to restart RStudio, you can do this - just make sure you save things if you have other stuff open.*

</br>

### Read data into R  

If you have downloaded the whole R project from the [CLE page](http://tiny.ucsf.edu/dsiregr) or [GitHub](https://github.com/UCSF-DSI/RegR) for this workshop, you should have the data and directories all set to go for reading the dataset. Just double click on the **RegR.Rproj** file, and there will be no need to set your working directory before reading the data below.  

***HRS variables***

```{r echo=FALSE}
Variable <- c("id", "female", "age_2014", "depress_dx", "bp_dx", "bp_rx", "bp_sys", "bp_dia",
              "polygen_map", "polygen_dep", "died", "mos_to_death")
Label <- c("unique patient identifiers (random)", "female (self-reported)", "age in 2014", "depression diagnosis (ever, self-reported)", "high blood pressure diagnosis (ever, self-reported)", "high blood pressure meds (current, self-reported)", "systolic blood pressure (measured)", "diastolic blood pressure (measured)", 
           "polygenic marker score ~ mean arterial pressure", "polygenic marker score ~ depression",
           "died (or censored)", "months to death (or censoring)")

hrsvars <- as.data.frame(cbind(Variable, Label))

hrsvars
```

We will first read the file called "hrs_analyze.csv" into R (calling this object simply `hrs`). Note: this assumes that you are in your main project directory. If you are not, please set your working didrectory to that.    

```{r}
hrs <- read.csv("../Data/hrs_analyze.csv")
hrs
```

### Inspect the data

```{r}
str(hrs) ## remember str stands for "structure" in R
head(hrs) 
```

</br>

### KNOWLEDGE CHECK: Read data into R 




