---
title: "Regression modeling with R Part 2"
subtitle: "Setup notebook"
author: "Karla Lindquist"
date: "March 31, 2022"
output: html_notebook
---

------------------------------------------------------------------------

### Create an RStudio project

I recommend downloading the complete project from [GitHub](https://github.com/UCSF-DSI/RegR) so that you don't have to worry about working directories. This is the main benefit of using RStudio projects - wherever you save them, everything will be relative to the .Rproj file.

If you want to creat an RStudio project from scratch, instructions are as follows:\
1. Go to the RStudio main menu, select **File -\> New Project**.\
2. Select **New Directory** 3. Select **Empty Project** 4. Give the project an informative **Directory Name** (this will be created for you). Use **Browse** to choose the location for this directory.\
*Optional*: check the "Create git repository" box if you know how to use git for version control. 5. Click **Create Project** button. This creates a new project directory, as well as a .Rproj file within it.

Within the project folder that you just created, create the following directories:

-   Data (save hrs_analyze.csv file from the Datasets section of the CLE page here)
-   Notebooks (if using R markdown)
-   Scripts (if creating new project from scratch)
-   Results (optional)

------------------------------------------------------------------------

### Start an R script

You can start a new R scipt using the RStudio drop-down menu that has a little page icon with a green **+** symbol (top left).

*Best practice tip*: At the top of your script, write comments that indicate:

```{r}
## Purpose: save commands from Regression Modeling with R Part 2
## Author: Karla Lindquist (karla.lindquist@ucsf.edu)
## Updated: March 31, 2022
```

You may also want to pre-define sections in your script. Using the `####` after a comment will create a navigable menu in RStudio (see bottom of script window).

```{r}
## Read data ####

## Kaplan-Meier and log rank tests ####

## Cox regression ####

```

------------------------------------------------------------------------

### Install and Load Required Packages

Next, install and load the packages that we will be using in all sections.

```{r}
pkgs <- c("tidyverse","jtools","survival","gplots","survminer") 

install.packages(pkgs, update = TRUE, ask = FALSE) 

for(package in pkgs) {
  do.call("library", list(package))
}
```

*Note: this will update packages if they already exist in your system. If asked to restart RStudio, you can do this - just make sure you save things if you have other stuff open.*

------------------------------------------------------------------------

### Read data into R

If you have downloaded the whole R project from [GitHub](https://github.com/UCSF-DSI/RegR) for this workshop, you should have the data and directories all set to go for reading the dataset. Just double click on the **RegR.Rproj** file, and there will be no need to set your working directory before reading the data below.

***HRS variables***

```{r echo=FALSE}
Variable <- c("id", "female", "age_2014", "depress_dx", "bp_dx", "bp_rx", "bp_sys", "bp_dia",
              "polygen_map", "polygen_dep", "died", "mos_to_death")
Label <- c("unique patient identifiers (random)", "female (self-reported)", "age in 2014", "depression diagnosis (ever, self-reported)", "high blood pressure diagnosis (ever, self-reported)", "high blood pressure meds (current, self-reported)", "systolic blood pressure (measured)", "diastolic blood pressure (measured)", 
           "polygenic marker score ~ mean arterial pressure", "polygenic marker score ~ depression",
           "died (or censored)", "months to death (or censoring)")

hrsvars <- as.data.frame(cbind(Variable, Label))

hrsvars
```

We will first read the file called "hrs_analyze.csv" into R (calling this object simply `hrs`). Note: this assumes that you are in your main project directory. If you are not, please set your working directory to that.

```{r}
hrs <- read.csv("../Data/hrs_analyze.csv")
```

------------------------------------------------------------------------

### Inspect the data

```{r}
glimpse(hrs) ## remember glimpse is similar to the str() or "structure" function in R
head(hrs) 
```
